---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r cache=TRUE}
unzip('activity.zip')
```

reading file
```{r cache=TRUE}
data <- read.csv('activity.csv')
head(data)
```
processing dates
```{r}
library(lubridate)
data$date <- ymd(data$date)
head(data)
```

## What is mean total number of steps taken per day?

Total steps by day
```{r}
library(tidyverse)
total_by_day <- data %>%
  group_by(date) %>%
  summarize(sum_steps_by_day = sum(steps, na.rm=T))
total_by_day
```
Histogram of steps by day
```{r}
total_by_day <- data %>%
  group_by(date) %>%
  summarize(sum_steps_by_day = sum(steps, na.rm=T))
total_by_day
hist(total_by_day$sum_steps_by_day)
```

Summary of data including means and medians
```{r}
summary(data)
```
Mean and median of steps by day ignoring NA values
```{r}
library(tidyverse)
table(data$steps)
mean_by_day <- data %>%
  group_by(date) %>%
  summarize(mean_steps_by_day = mean(steps, na.rm=T), median_steps_by_day= median(steps, na.rm=T))
mean_by_day
```

## What is the average daily activity pattern?

Time series plot of mean by day
```{r}
steps_by_interval <- data %>%
  group_by(interval) %>%
  summarize(mean_steps_by_interval = mean(steps, na.rm=T), median_steps_by_interval = median(steps, na.rm=T))

with(steps_by_interval, plot(interval, mean_steps_by_interval, type='l'))
```

5 minute interval that contains the max number of steps
```{r}
steps_by_interval[which.max(steps_by_interval$mean_steps_by_interval),]
```

## Imputing missing values

Imputing missing with Median with Caret
```{r}
library(caret)
data_imp_model = preProcess(data, "medianImpute")
idata = predict(data_imp_model, data)
summary(idata)
```

Histogram of steps by day
```{r}
total_by_day <- idata %>%
  group_by(date) %>%
  summarize(sum_steps_by_day = sum(steps, na.rm=T))
total_by_day
hist(total_by_day$sum_steps_by_day)
```
#Mean and median of the imputed data; mean is lower, median is the same
```{r}
summary(idata$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
find_weekend <- function(day_of_week){
  if(day_of_week == "Saturday"||day_of_week == "Sunday"){
    paste('Weekend')
  }
  else{
    paste('Weekday')
  }
}
idata <- idata %>% mutate(day_of_week = weekdays(date), isWeekend = ifelse(day_of_week == "Saturday" | day_of_week == "Sunday", "Weekend", "Weekday")) %>% mutate(isWeekend = as.factor(isWeekend))
summary(idata)
table(idata$day_of_week)
table(idata$isWeekend)
```
Making plot
```{r}
library(ggplot2)
isteps_by_interval <- idata %>%
  group_by(interval, isWeekend) %>%
  summarize(mean_steps_by_interval = mean(steps), median_steps_by_interval = median(steps))
isteps_by_interval

ggplot(data=isteps_by_interval, aes(x=interval, y=mean_steps_by_interval)) + geom_line() + facet_grid(isWeekend ~ .)
```



